What We Eval in the Shadows

### 2021-08-09 paper/main.tex
More Kaggle removal

\paragraph{Consistency} It is interesting how many different
forms any given site sees. The more forms, the harder it is to characterize
behavior at that site. Luckily, \packageNbOneMinimizedPercent of sites only see
a single form. In other words, a compiler could correctly predict the
\emph{shape} of an expression by looking just at the first call in
\packageNbOneMinimizedPercent of the sites. Few sites are highly polymorphic (8
or more different forms). These include the pipe operator of the \c{magrittr}
package, used to compose functions.

\paragraph{Discussion} The variety of uses of \eval is evidenced by the number
of different forms. In comparison, JavaScript \eval usage was more
straightforward and more predictable as reported by \citet{oopsla12b}, with
98.7\% of the sites with only one form. Nevertheless, simple forms dominate in R
also, and there are many cases where \eval could be replaced by less powerful
constructs.

@@

Strings could correlate with dynamic code loading. This is what the base
functions \c{source} and \c{sys.source} do. We observed few calls
(\packageNbParseFromFileSites in total) that consume the result of calling
\c{parse} on a file. Most of the calls build strings programmatically. We also
identified one function \c{invokeRestartInteractively} that prompts the user for
input, parses it, and passes it to \eval. The use of strings seems to correlate
with less sophisticated programmers; in Kaggle,
\kaggleParseExprsSitePercent of sites use strings.

@@

\Eval may peform side-effects. From a compiler's perspective, we care about
effects that can be observed --- \ie variable definitions, updates, and removals
that are visible after a call finishes. Knowing in which environments these
side-effects happen, can help us determine how much of the compiler knowledge
about the program will be invalidated. Our analysis records information about
every environment. From the recorded data, we discard side-effects coming from
unit testing frameworks.\footnote{In the corpus, we have \c{RUnit, testthat,
  tinytest} and \c{unitizer} testing frameworks.} They run their tests via
\eval, and thus everything becomes a side-effect.

>>>>

\paragraph{Consistency} It is interesting how many different forms any given
site sees. The more forms, the harder it is to characterize behavior at that
site. Luckily, \packageNbOneMinimizedPercent of sites only see a single form.
In other words, a compiler could correctly predict the \emph{shape} of an
expression by looking just at the first call in \packageNbOneMinimizedPercent
of the sites. Few sites are highly polymorphic (8 or more different forms).
These include the pipe operator of the \c{magrittr} package, used to compose
functions, \eg instead of \c{f(g(h(x),y),z)} one can write \c{h(x) \%>\% g(y) \%>\%
f(z)}.

\paragraph{Discussion} The variety of uses of \eval is evidenced by the number
of different forms. In comparison, JavaScript \eval usage was more
straightforward and more predictable as reported by \citet{oopsla12b}, with
98.7\% of the sites with only one form. Nevertheless, simple forms dominate
in R. Among these there are many cases where \eval could be replaced by less
powerful constructs such as \c{get} to lookup a name in an environment,
\c{assign} to set a value to a name in an environment or \c{call} to
reflectively call a function.
% TODO: how many receive just the simple form

@@

Strings could correlate with dynamic code loading. This is what the base
functions \c{source} and \c{sys.source} do. We observed few calls
(\packageNbParseFromFileSites in total) that consume the result of calling
\c{parse} on a file. Most of the calls build strings programmatically. We also
identified one function \c{invokeRestartInteractively} that prompts the user for
input, parses it, and passes it to \eval.

@@

\Eval may perform side-effects. From a compiler's perspective, we care about
effects that can be observed --- \ie variable definitions, updates, and removals
that are visible after a call finishes. Knowing in which environments these
side-effects happen, can help us determine how much of the compiler knowledge
about the program will be invalidated. Our analysis records information about
every environment. From the recorded data, we discard side-effects coming from
unit testing frameworks.\footnote{In the corpus, we have \c{RUnit, testthat,
  tinytest} and \c{unitizer} testing frameworks.} They run their tests via
\eval, and thus everything becomes a side-effect.
